---
title: "test-grades"
author: "Kenza REDWANE"
format: 
  html: 
    code-fold: true
---

### Introduction

## 1.1 Study organisation

## 1.2 Data set

```{r}
#| message: false
#| cache: false
here::i_am("grades(graded).csv")
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
```

Question 1: Load data
```{r}
library(readr)
grades <- read_csv("/Users/macbookpro/Desktop/Master/Semester One/Data Management and Programming/test-grades/grades(graded).csv")

```
## Simple analyses:

Question 2:
```{r}
num_grades <- nrow(grades)
paste("The data set contains", num_grades, "grades")
```

Question 3:
```{r}
num_students <- grades |> distinct(id) |> nrow()
paste("The data set contains", num_students, "students")
```

Question 4:
```{r}
students <- grades |> distinct(id, group)
students |> slice_head(n = 5)
```

Question 5: 
```{r}
students_per_group <- students |> count(group)
ggplot(students_per_group, aes(x = factor(group), y = n)) +
  geom_col(fill='sky blue')
```
Question 6:
```{r}
arch_sculpture_avg <- grades |> 
  filter(course == "Architecture and Sculpture") |> 
  group_by(group) |> 
  summarise(avg_grade = mean(grade, na.rm = TRUE))

ggplot(arch_sculpture_avg, aes(x = factor(group), y = avg_grade)) +
  geom_col(fill="steel blue") 
```
we can compute another graphical representation of the Average architecture
```{r}
average_architecture <- grades %>%
  filter(course == "Architecture and Sculpture") %>%
  group_by(group) %>%
  summarise(average_grade = mean(grade, na.rm = TRUE), .groups = 'drop')
ggplot(average_architecture, aes(x = group, y = average_grade)) +
  geom_point(size = 3, color = "blue") +
  geom_line(aes(group = 1), color = "blue", linetype = "dashed") +
  theme_minimal()
```
Question 7: Comparison
```{r}
ggplot(grades, aes(x = grade, fill = factor(semester))) +
  geom_density(alpha = 0.5) +
  labs(fill = "Semester")
```
### Attendance analysis

Question 8:
```{r}
grades_per_student <- grades |> 
  group_by(id, group) |> 
  summarise(num_grades = n())

summary_stats <- grades_per_student |> 
  summarise(min_grades = min(num_grades),
  max_grades = max(num_grades),
  avg_grades = mean(num_grades),
  med_grades = median(num_grades))
summary_stats
```

Question 9
```{r}
martial_grades <- grades |> 
  filter(course == "Martial Arts and Self-Defense") |> 
  group_by(id, group) |> 
  summarise(num_martial_grades = n())
martial_grades |> slice_head(n = 5)
```
Question 10:
```{r}
distribution_martial_grades <- martial_grades |> 
  count(num_martial_grades)
ggplot(distribution_martial_grades, aes(x = num_martial_grades, y = n)) +
  geom_col(fill ='darkred')
  
```

Question 11:
```{r}
ggplot(martial_grades, aes(x = factor(group), y = num_martial_grades)) +
  geom_boxplot() 

```

## Grade Analysis

Question 12: 
```{r}
avg_grades_per_course <- grades |> 
  group_by(id, group, course) |> 
  summarise(avg_grade = mean(grade, na.rm = TRUE)) |> 
  ungroup() |>
  pivot_wider(names_from = course, values_from = avg_grade)

avg_grades_per_course |> 
  select(id, group, starts_with("Agriculture"), starts_with("Architecture")) |> 
  slice_head(n = 5)
```

Question 13: 
```{r}
ggplot(avg_grades_per_course, aes(x = `Traditional Medicine and Herbalism`, y = `Music and Dance`)) +
  geom_point(color = "blue", alpha = 0.6) +
  theme_minimal()
```

Question 14: Correlation between the average grades in astrology and divination & Architecture and Sculpture
```{r}
correlation_by_group <- avg_grades_per_course |>
  group_by(group) |>
  summarise(correlation = cor(`Astrology and Divination`, `Architecture and Sculpture`, use = "complete.obs"))

correlation_by_group
```

Question 15: 
```{r}
strongest_corr_group <- correlation_by_group |> 
  slice_max(abs(correlation), n = 1) |> 
  pull(group)
avg_grades_per_course |>
  filter(group == strongest_corr_group) |>
  ggplot(aes(x = `Astrology and Divination`, y = `Architecture and Sculpture`)) +
  geom_point(color = "red") +
  theme_minimal()
```

Question 16: 
```{r}
final_grades <- avg_grades_per_course |>
  rowwise() |> 
  mutate(final_grade = mean(c_across(starts_with("Agriculture"):`Traditional Medicine and Herbalism`), na.rm = TRUE)) |>
  ungroup() |>
  select(id, group, final_grade) |>
  arrange(desc(final_grade))

final_grades |> slice_head(n = 5)
```

Question 17: Study differences in final grades between groups
```{r}
ggplot(final_grades, aes(x = factor(group), y = final_grade)) +
  geom_boxplot(fill = "lightblue") +
  theme_minimal()
```

Question 18: 
```{r}
library(dplyr)
library(purrr)
final_results <- grades %>%
  group_by(id, group) %>%
  summarise(final_grade = mean(grade, na.rm = TRUE),
            semester_averages = list(mean(grade)),  
            .groups = 'drop'  ) %>%
  mutate( average_semester = map_dbl(semester_averages, mean, na.rm = TRUE), 
    pass = final_grade >= 5 & average_semester >= 10) %>%
  select(id, group, final_grade, pass)  
final_results
```

Question 19: 
```{r}
non_passing_high_grades <- final_results %>%
  filter(pass == FALSE) %>%
  nrow()
non_passing_high_grades
```

Question 20: 
```{r}
pass_rate_per_group <- final_results %>%
  group_by(group) %>%
  summarise(pass_rate = mean(pass) * 100, .groups = 'drop')  
pass_rate_per_group

ggplot(pass_rate_per_group, aes(x = group, y = pass_rate)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal()
```
